\yourname

\activitytitle{Background  and syllabus questions -- Math 3280}{Do your best with these questions and turn this sheet in on Thursday.
Some of them reference information from the syllabus. This assignment is worth 10 points}

\blist{0.8in}

\item (2 points) What mathematics courses have you already taken in college?
It's OK to just list the numbers, like Math 3410.

\item (2 points) Please list all the courses you are taking this semester, aside from this one.

\item (2 points) Including this semester, how many semesters have you been at BGSU?

\item How comfortable are you with the ``definition, example, theorem, proof'' progression in mathematics classes?

\item What kinds of experiences have you had in the past with proofs?

\item Have you ever had success reading a mathematics textbook and really learning from it?  If so, please tell what course and what made it work.  If not, please tell me what you think prevented you from being able to read the book.

\item (2 points) If you take the elevator to the fourth floor, do you turn right or left to get to my office?

\item (2 points) What is the most interesting thing to you on the door of my office?

\item Do you have a hard copy of the textbook that you can read?  Electronic copy?

\item Do you have any interest in going to graduate school?  Please explain.

\item Do you have any questions or concerns about the coursework?

\item Do you have any questions or concerns about the grading?

\item What is the most likely reason that you will miss class?  I'm just curious.

\item Please let me know anything you think I should know about you.  I'll read it all.  Sometimes people like to tell about their hobbies, movies they like, other academic interests, clubs they're in, where they're from, etc.

\elist

\vfill          % pad the rest of the page with white space
