\activitytitle{Reading assignment, Chapter 18}{Reading due on November 30, notes and exercises due on December 2.}

Read and understand Chapter 18 of the textbook by Daepp and Gorkin, called ``Mathematical Induction,'' up to the statement, but not the proof, of Theorem 18.6.

Mathematical induction and recursion play an important role especially in discrete mathematics. Prepare to move slowly and think carefully. To understand the proof of Theorem 18.1, you will need {\bf Well-ordering principle of the natural numbers}: Every nonempty subset of the natural numbers contains a minimum.

Read Theorem 18.1 and then {\bf do Problem 18.1} and {\bf Problem 18.3}. Follow the steps in Theorem 18.1, defining the assertion $P(n)$ for the problem first. You will need the condition ``$P(n)$ is true'' to show the induction step. Work through Exercises 18.3 to 18.5 and then {\bf do Problem 18.9} without going back to Exercise 18.5. You can do it!

Recursion is a very useful tool to define functions, sequences and sets. Before you move to Theorem 18.6, read the definition of $n$ factorial for $n\in \mathbb{N}$. Write out $3!, 4!$ and $5!$.  As an exercise, simplify $\frac{6!}{2!4!}$. More generally, simplify $\frac{n!}{m!(n-m)!}$ where $n$ and $m$ are two positive integers with $n\geq m$. These fractions are called {\em binomial coefficients} and are useful in probability.

Here is another example of using recursion:  
Let $n\in \mathbb{Z^+}$. Consider the function $S(n) = S(n-1) + n$ with $S(0) = 0$. Write out $S(1), S(2), S(3)$ and $S(4)$. Can you figure out what this function does for us?  Together with Problem 18.1, you should be able to see the connection between induction and recursion.

Theorem 18.6 shows the existence and uniqueness of a recursive function $g: N \rightarrow X$ given a function $f: X \rightarrow X$ and $a\in X$, where $X$ is a nonempty set. The function $g$ satisfies
\begin{itemize}
\item[(i)] The base step: $g(0) = a$, and
\item[(ii)] The recursive step: $g(n+1)=f(g(n))$ for all $n \in \mathbb{N}$.
\end{itemize}

\noindent The proof of Theorem 18.6 is too long for us to read this semester.  You can come back it later.

\noindent As with the previous chapters,
\blist{0.0in}
\item Read somewhere quiet, minimizing distractions from phones and friends
\item Note the time that you start and stop reading, and add up the minutes
\item Read with a pencil in your hand and your notebook open in front of you
\item Write a sentence to summarize each paragraph, re-draw diagrams, work out examples and exercises on your own
\item Look up words you don't know, and write down ones you really don't know
\item Read slowly.
\item At the end, tally up how much time you have spent on reading this chapter.
Write this number in your notebook and remember the number when you come to class.
\elist

\vfill          % pad the rest of the page with white space
