\activitytitle{Reading assignment Chapter 8}{Due on Tuesday, October 24.  20 points}

Read Chapter 8 in the book by Daniel Solow.
This chapter is about negation of logical statements, especially of statements containing quantifiers.
This just takes some practice and you'll be good at it.
The key text is steps 1, 2, 3 at the top of page 95.


\vspace{0.1in}
\noindent
{\bf Specific requirements}
\vspace*{-0.15in}

\begin{itemize}
\item Write the NOT of this statement:  ``For all $x \in \R$, $\ln(x) < 14$.

\item Write the NOT of this statement:  ``For all $a \in [2,5]$, there exists $b \in [2,5]$ such that $a < b$.

\item Write the NOT of this statement:  ``There exists $a \in A$ such that for all $b \in B$, $f(b) < a$.

\item Write the NOT of this statement:  ``For all $\varepsilon > 0$, there exists $\delta > 0$ such that for all $x$ with $|x-a| < \delta$, $|f(x) - L| < \varepsilon$.

\item Write the NOT of this statement:  ``For all $a \in \R$, for all $\varepsilon > 0$, there exists $\delta > 0$ such that for all $x$ with $|x-a| < \delta$, $|f(x) - f(a)| < \varepsilon$.

\item Do exercise 8.2.  Note that  taking the NOT of ``For all a and for all b, something happens'' becomes ``There exists a and there exists b such that, NOT something happens.''

\item Do exercise 8.3.  For part (a), note that ``There is no integer $n$ with ...'' is how you write English for the logical statement ``NOT (there exists an integer $n$ with ...)''.  So it's easy to negate.  For all parts, note that you are just going to put NOT in front of the bold word, and then write the NOT of what comes after ``if and only if''.

\item Do exercise 8.7.  Clearly identify the logical statements $A$ and $B$ in each case, and then clearly state NOT $B$ and NOT $A$.  Leave out statements such as $k$ is an integer; these are like the fabric of reality, not to be negated in these exercises.  Make clear what you will work forward from and what you will work backward from.  Note that you do not have to do the proofs, but if you see how to do them, you might as well do them.

\item Is this statement true?  ``For all $a \in [2,5]$, there exists $b \in [2,5]$ such that $a < b$.  If so, explain.  If not, find a counterexample.

\item Is this statement true?  ``For all $a \in (2,5)$, there exists $b \in (2,5)$ such that $a < b$.  If so, explain.  If not, find a counterexample.

\item At the end, tally up how much time you have spent on this chapter.
Write this number in your notebook.
Bring your notebook to class and turn it in for grading.
\end{itemize}

\noindent
{\bf General comments}

Set yourself up in a place where you won't be disturbed.
Read slowly, and write notes in your own words that reflect your understanding of the material.
