\yourname

\activitytitle{Sum and dot product of 3--dimensional vectors}{\vspace*{-0.2in}}

\overview{In Calculus III and Linear Algebra, we define vectors and work with them.  
They have a geometric interpretation, but here we will simply give an algebraic definition of 3--dimensional vectors and some operations on them and work with their algebraic properties.
This activity illustrates proofs in which all that is needed is the definition and a ``rewrite'' proof, where you can work forward and backward to show a series of equalities.
Notice how we often use the same definition twice in one proof, once to ``unpack'' and the second time to ``re--pack.''}

\definition{3--dimensional vector\label{3dvectordef}}{A three--dimensional vector is an ordered triple $\vect{ a_1, a_2, a_3 }$, where $a_1, a_2,$ and $a_3$ are real numbers.
The numbers  $a_1, a_2,$ and $a_3$ are called {\em components} of the vector.
}

\notation{A 3--dimensional vector $\vect{ a_1, a_2, a_3 }$ is often denoted by a single letter with an arrow over the top, like this $\vec{a}$.  When it is written like $\vect{ a_1, a_2, a_3 }$ it is said to be in {\em open form.}
The commas and brackets are part of the definition and are important.}

\definition{Equality of 3--dimensional vectors\label{3dvectorequalitydef}}{3--dimensional vectors $\vect{ a_1, a_2, a_3 }$ and $\vect{ b_1, b_2, b_3 }$ are equal if $a_1=b_1, a_2=b_2,$ and $a_3 = b_3$.  Note: The order of the numbers is important.}

\definition{Sum of 3--dimensional vectors\label{3dvectorsumdef}}{The sum of 3--dimensional vectors $\vect{ a_1, a_2, a_3 }$ and $\vect{ b_1, b_2, b_3 }$ is the 3--dimensional vector $\vect{ a_1+b_1, a_2+b_2, a_3+b_3 }$.  We write $\vec{a} \oplus \vec{b}$ for the sum of $\vec{a}$ and $\vec{b}$, using a new symbol so we don't confuse addition of vectors with addition of real numbers.}

\example{Is $\vect{ 3,9,12 }$ a 3--dimensional vector? Explain.  

\vspace*{0.1in}
\noindent
Is it equal to $\vect{12,3,9}$?  Explain.}{0.1in}

\example{Is $\vect{ \sqrt{3},\sqrt[3]{9},\sqrt{-12} }$ a 3--dimensional vector? Explain.}{0.3in}

\example{Is $\vect{ 8,13.35321,\pi,-7 }$ a 3--dimensional vector? Explain.}{0.3in}

\example{Is $\vect{ 3 + 9 + 12 }$ a 3--dimensional vector? Explain.}{0.3in}

\example{Is $\vect{ \left[ \begin{array}{cc} 6 & 0 \\ 2 & 5 \end{array} \right], -4, 7 }$ a 3--dimensional vector? Explain.}{0.4in}

\example{Let $x$ be a real number.  Is $\vect{ \frac{14}{3},2-7x,\sqrt{16} }$ a 3--dimensional vector? Explain.}{0.3in}

\stop{Compare your answers to the questions above with the members of your group.  Make sure you agree on everything.}
\pagebreak

\example{Calculate the sum of $\vec{c} = \vect{ 12,-5,3 }$ and $\vec{d} = \vect{ 6,4,-11 }$.  Start by writing $\vec{c} \oplus \vec{d} = \ldots$ and write the vectors in open form next.


\vspace*{0.2in}
\noindent
Calculate $\vec{d} \oplus \vec{c}$ in the same way.
}{0.1in}

\show{You are going to show that addition of 3--dimensional vectors is commutative.  
Fill in the blanks.
This is a ``rewrite'' proof.
You can work forward from the top, backward from the bottom, or a bit of both.\\
Let $\vec{a}$ and $\vec{b}$ be 3--dimensional vectors.  Then,
\begingroup
\addtolength{\jot}{0.7em}
\begin{eqnarray*}
    \vec{a} \oplus \vec{b}
    &=& \vect{ \qqq,\qqq,\qqq } \oplus \vect{\qqq,\qqq,\qqq} \qqqq \qqqq \qqqq \qqqq \\
    &=& \vect{ \qqq\qqq,\qqq\qqq,\qqq\qqq } \\
    &=& \vect{ \qqq\qqq,\qqq\qqq,\qqq\qqq } \\
    &=& \vect{ \qqq,\qqq,\qqq } \oplus \vect{\qqq,\qqq,\qqq} \\
    &=& \vec{b} \oplus \vec{a}
\end{eqnarray*}
\endgroup
We have seen that $\vec{a} \oplus \vec{b} = \vec{b} \oplus \vec{a}$.
We made no further assumption about $\vec{a}$ and $\vec{b}$.
Thus, for all 3--dimensional vectors $\vec{a}$ and $\vec{b}$, we know that  $\vec{a} \oplus \vec{b} = \vec{b} \oplus \vec{a}$.
Thus, addition of 3--dimensional vectors is commutative.
}{0in}

\show{Go back to each line of the proof above and give exactly one reason for the equality on that line at the very right side of the line.  
The first one is ``Write in open form.''  
Two of them are Definition \ref{3dvectorsumdef}.
In the middle you will use the fact that addition of real numbers is commutative.  
Thus, at the heart of it, commutativity of vector addition comes from commutativity of addition of real numbers.}{0in}

\show{Show that addition of 3--dimensional vectors is associative.
Start with arbitrary 3--dimensional vectors $\vec{a}, \vec{b},$ and $\vec{c}$.
Write $(\vec{a} \oplus \vec{b}) \oplus \vec{c}$ and rewrite it until it becomes $\vec{a} \oplus (\vec{b} \oplus \vec{c})$.
Take small steps and write exactly one reason for each equality.
Since you know what equality you need to show, you can work forward from the top, backward from the bottom, or both.\\
Let $\vec{a}, \vec{b},$ and $\vec{c}$ be \blank{3in}.
\begingroup
\addtolength{\jot}{0.7em}
\begin{eqnarray*}
    (\vec{a} \oplus \vec{b}) \oplus \vec{c}
    &=&  \qqqq \qqqq \qqqq \qqqq \qqqq\qqqq\qqqq\qqqq \qqqq \\
    &=&  \\
    &=&  \vect{ ( \qqq + \qqq ) + \qqq, ( \qqq + \qqq ) + \qqq, ( \qqq + \qqq ) + \qqq} \\
    &=&  \vect{ \qqq + ( \qqq + \qqq), \qqq + ( \qqq + \qqq), \qqq + ( \qqq + \qqq)} \\
    &=&  \\
    &=&  \\
    &=& \vec{a} \oplus (\vec{b} \oplus \vec{c})
\end{eqnarray*}
\endgroup
We have seen that ... 
}{0in}

\stop{Compare your argument to the rest of the members of your group.  Make sure that you agree on absolutely every step and every justification.}
\pagebreak

\definition{Scalar product for 3--dimensional vectors}{Let $c$ be a real number and let $\vec{a} = \tvec{a}$ be a 3--dimensional vector.  The {\em scalar product} of $c$ and $\vec{a}$ is a 3--dimensional vector defined as:
\[
    c\vec{a} = \vect{ca_1,ca_2,ca_3}.
\]}

\example{Let $c = 3$ and $\vec{a} = \vect{7,-4,\sqrt{2}}$.  Calculate $c\vec{a}$, starting by writing $c\vec{a} = 3 \vect{7, -4,\sqrt{2}} = \ldots$.}{0.3in}

\example{Calculate $\pi \vect{9,4,1}$ = }{0.2in}

\example{Calculate $(2+\sqrt{3}) \vect{5,b,c}$ = }{0.2in}

\show{You are going to show that the scalar product is distributive over vector addition.  First use the word ``Let'' to settle on one real number $c$ and two 3--dimensional vectors, $\vec{a}$ and $\vec{b}$.  
Then start with the expression $c(\vec{a} \oplus \vec{b})$ and rewrite it three times.
Then, move to the last expression and work backwards, until you meet in the middle.  
Provide one reason for each equality, on the right, on the same line as the equality.
At the end, follow the model to conclude that you have shown distributivity in general.

\vspace{0.2in}
\noindent
Let ...
\begingroup
\addtolength{\jot}{0.7em}
\begin{eqnarray*}
    c(\vec{a} \oplus \vec{b})
    &=&  \qqqq \qqqq \qqqq \qqqq \qqqq\qqqq\qqqq\qqqq \qqqq \\
    &&  \\
    &&  \\
    &&  \\
    &&  \\
    &&  \\
    &&  \\
    &&  \\
    &&  \\
    &=& c\vec{a} \oplus c\vec{b}
\end{eqnarray*}
\endgroup
We have seen that ...
}{0.5in}

\vfill

\stop{Check over what everyone in your group has done, and make sure that you completely agree.}

\newpage

\show{Show that the scalar product is distributive over real number addition.  
Start with ``Let.''  
Write $(c+d)\vec{a}$ and rewrite it until it equals $c\vec{a} \oplus d\vec{a}$.
Work forward from the top and backward from the bottom.
Provide one reason for each equality.
At the end, follow the model to conclude that this shows distributivity in general.
Explain why some addition signs are $+$ and others are $\oplus$.}{4in}

\definition{Zero vector}{The vector $\vect{0,0,0}$ is a special 3--dimensional vector, called the {\em zero vector}.
We denote it by $\vec{0}$.}

\definition{Additive inverse}{Let $\vec{a}$ be a 3--dimensional vector, with open form $\tvec{a}$.
Define a new vector by $-\vec{a} = \vect{-a_1,-a_2,-a_3}.$
It is called the {\em additive inverse} of $\vec{a}$.}

\show{Let $\vec{a}$ be a 3--dimensional vector.  
Use a rewrite proof to show that $\vec{a} \oplus \vec{0} = \vec{a}$.  
This is called the {\em additive identity} property.
It's not very exciting.  Make a general conclusion.}{1.2in}

\show{Let $\vec{a}$ be a 3--dimensional vector, and let $-\vec{a}$ be its additive inverse.  
Use good form to show that $\vec{a} \oplus (-\vec{a}) = \vec{0}$.  
This is called the {\em additive inverse} property.
This is also not very exciting.  Make a general conclusion.}{0in}

\pagebreak
\definition{Dot product of 3--dimensional vectors}{The dot product of 3--dimensional vectors $\vect{ a_1, a_2, a_3 }$ and $\vect{ b_1, b_2, b_3 }$ is the real number $a_1 b_1 + a_2 b_2 + a_3 b_3$.}

\notation{The dot product of 3--dimensional vectors $\vec{a}$ and $\vec{b}$ is denoted $\vec{a} \bullet \vec{b}$.}

\example{Calculate the dot product of $\vec{a} = \vect{ 12,-5,3 }$ and $\vec{b} = \vect{ 6,4,-11 }$  Do this by writing
\begin{eqnarray*}
    \vec{a} \bullet \vec{b} &=& \tvec{a} \bullet \tvec{b}\\
    &=& a_1 b_1 + a_2 b_2 + a_3 b_3
\end{eqnarray*}
and then substituting in the numbers.
This makes the calculation just a matter of rewriting, so it is a good way to do calculations like this.}{1.3in}

\show{Show that the dot product is commutative, just as multiplication of real numbers is commutative.
Start with ``Let''.
Write one expression at the top of the space below, and write your goal expression at the bottom, and then work
forward and backward until you have a rewrite proof.
Follow the models from previous examples, and be sure to make a general conclusion.}{3.5in}

\example{Calculate $\vec{a} \bullet \vec{0}$.  Is this a general result?  If so, make your calculation into a general result.}{0in}

\newpage

\show{Show that the dot product is distributive over vector addition.
That is, show that $(\vec{a} \oplus \vec{b}) \bullet \vec{c} = \vec{a} \bullet \vec{c} + \vec{b} \bullet \vec{c}.$  
Start with ``Let''.  
Write the first expression at the top, the last expression at the bottom, and then work forward and backward.
Also explain why one addition symbol is $\oplus$ and the other is $+$.}{3.5in}

\show{Let $\vec{a}$ and $\vec{b}$ be 3--dimensional vectors and let $c$ be a real number.  Show in general that $c(\vec{a} \bullet \vec{b}) = (c\vec{a}) \bullet \vec{b} = \vec{a} \bullet (c\vec{b})$.  Since there are two equalities to show, think about how you will organize the equalities.}{0in}

\vfill          % pad the rest of the page with white space
